---
title: "BACS HW12 - 109006234"
date: "May 7th 2023"
output:
  pdf_document:
    latex_engine: xelatex
author: "Credit: 109006278"
geometry: margin = 0.8in
---
# Loading the data
```{r}
cars <- read.table("auto-data.txt", header=FALSE, na.strings = "?")
names(cars) <- c("mpg", "cylinders", "displacement", 
                "horsepower", "weight", "acceleration", 
                "model_year", "origin", "car_name")
vars <- c("mpg", "weight", "acceleration", 
        "model_year", "origin", "cylinders")
cars <- cars[vars]
cars_log <- with(cars, data.frame(log(mpg), log(weight), 
            log(acceleration), model_year, origin, log(cylinders)))
```

# Problem 1
## (a) Let’s visualize how weight might moderate the relationship between acceleration and mpg:
```{r}
weight_mean_log <- log(mean(cars$weight))
```
**(i) Create two subsets of your data, one for light-weight cars (less than mean weight) and one for heavy cars (higher than the mean weight)**
```{r}
#Light Cars Regression
light_log <- subset(cars_log, log.weight. < weight_mean_log)
light_reg <- with(light_log, lm(log.mpg. ~ log.acceleration.))

#Heavy Cars Regression
heavy_log <- subset(cars_log, log.weight. >= weight_mean_log)
heavy_reg <- with(heavy_log, lm(log.mpg. ~ log.acceleration.))
```
**(ii) Create a single scatter plot of acceleration vs. mpg, with different colors and/or shapes for light versus heavy cars**
```{r, out.width="80%", fig.align = "center"}
with(light_log, plot(log.acceleration., log.mpg., pch=1, 
        col="darkgreen", main = "Acceleration vs. MPG"))
with(heavy_log, points(log.acceleration., log.mpg., pch=5, col="blue"))
```
**(iii) Draw two slopes of acceleration-vs-mpg over the scatter plot: one slope for light cars and one slope for heavy cars**
```{r, out.width="80%", fig.align = "center"}
with(light_log, plot(log.acceleration., log.mpg., pch=1, 
        col="darkgreen", main = "Acceleration vs. MPG"))
with(heavy_log, points(log.acceleration., log.mpg., pch=5, col="blue"))
abline(light_reg, col="darkgreen", lwd=2)
abline(heavy_reg, col="blue", lwd=2)
```

## (b) Report the full summaries of two separate regressions for light and heavy cars where log.mpg. is dependent on log.weight., log.acceleration., model_year and origin
**Light Cars**
```{r}
regr_light_full <- lm(log.mpg.~ log.weight. + log.acceleration. + 
                    model_year + factor(origin), data=light_log)
summary(regr_light_full)
```
**Heavy Cars**
```{r}
regr_heavy_full <- lm(log.mpg.~ log.weight. + log.acceleration. +   
                    model_year + factor(origin), data=heavy_log)
summary(regr_heavy_full)
```

## (c) Using your intuition only: What do you observe about light versus heavy cars so far?
From the plot above, we can confirm that there are more light cars compared to heavy cars, as there are more dark-green data points. I think that the slope of heavy cats are more fitted compared to the other data available. 

# Problem 2
## (a) Considering weight and acceleration, use your intuition and experience to state which of the two variables might be a moderating versus independent variable, in affecting mileage.
In my opinion, acceleration might be a moderating vs. independent variable, while at the same time it will affect mpg.

## (b) Use various regression models to model the possible moderation on log.mpg.
**(i) Report a regression without any interaction terms**
```{r}
no_inter <- lm(log.mpg. ~ log.weight. + 
              log.acceleration. + model_year + 
              factor(origin), data=cars_log)
summary(no_inter)
```
**(ii) Report a regression with an interaction between weight and acceleration**
```{r}
weight_acc_inter <- lm(log.mpg. ~ log.weight. + log.acceleration. + 
                    log.weight. * log.acceleration., data=cars_log)
summary(weight_acc_inter)
```
**(iii) Report a regression with a mean-centered interaction term**
```{r}
mean_weight <- scale(cars_log$log.weight., center = TRUE, scale = FALSE)
mean_acc <- scale(cars_log$log.acceleration., center = TRUE, scale = FALSE)
mean_mpg <- scale(cars_log$log.mpg., center = TRUE, scale = FALSE)

mean_reg <- lm(mean_mpg ~ mean_acc + mean_weight + mean_acc*mean_weight)
summary(mean_reg)
```
**(iv) Report a regression with an orthogonalized interaction term**
```{r}
dot_product <- cars_log$log.weight. * cars_log$log.acceleration.
inter_reg <- lm(dot_product ~ 
                cars_log$log.weight. + cars_log$log.acceleration.)
inter_ortho <- inter_reg$residuals
ortho_inter <- lm(log.mpg. ~ log.weight. + log.acceleration. + 
                inter_ortho, data=cars_log)
summary(ortho_inter)
```
## (c) For each of the interaction term strategies above (raw, mean-centered, orthogonalized) what is the correlation between that interaction term and the two variables that you multiplied together?
```{r}
#Raw
a <- cor(cars_log$log.weight.*cars_log$log.acceleration., cars_log$log.weight.)
b <- cor(cars_log$log.weight.*cars_log$log.acceleration., cars_log$log.acceleration.)

#Mean-centered
c <- as.vector(cor(mean_acc*mean_weight, mean_weight))
d <- as.vector(cor(mean_acc*mean_weight, mean_acc))

#Orthogonalized
e <- cor(inter_reg$residuals, cars_log$log.weight.)
f <- cor(inter_reg$residuals, cars_log$log.acceleration.)

cor_mat <- matrix(c(a,b,c,d,e,f), ncol=2,byrow=TRUE)
rownames(cor_mat) <- c("raw", "mean-centered", "orthogonalized")
colnames(cor_mat) <- c("log.weight.","log.acceleration")
round(cor_mat,2)
```
# Problem 3
**Let’s check whether weight mediates the relationship between cylinders and mpg, even when other factors are controlled for.  Use log.mpg., log.weight., and log.cylinders as your main variables, and keep log.acceleration., model_year, and origin as control variables**

## (a) Let’s try computing the direct effects first:
**(i) Model 1: Regress log.weight. over log.cylinders. only**
```{r}
reg_weight_cyl <- lm(log.weight.~ log.cylinders., data=cars_log)
summary(reg_weight_cyl)
```
**(ii) Model 2: Regress log.mpg. over log.weight. and all control variables**
```{r}
regr_mpg_weight <- lm(log.mpg.~ log.weight., data=cars_log)
summary(regr_mpg_weight)
```
## (b) What is the indirect effect of cylinders on mpg?
```{r}
reg_weight_cyl$coefficients[2] * regr_mpg_weight$coefficients[2]
```
## (c) Let’s bootstrap for the confidence interval of the indirect effect of cylinders on mpg
```{r}
boot_mediation <- function(model1, model2, dataset) {
  boot_index <- sample(1:nrow(dataset), replace=TRUE)
  data_boot <- dataset[boot_index, ]
  regr1 <- lm(model1, data_boot)
  regr2 <- lm(model2, data_boot)
  return(regr1$coefficients[2] * regr2$coefficients[2])
}
```
**(i) Bootstrap regression models 1 & 2, and compute the indirect effect each time: What is its 95% CI of the indirect effect of log.cylinders. on log.mpg.?**
```{r}
set.seed(42)
indirect <- replicate(2000, 
        boot_mediation(reg_weight_cyl, regr_mpg_weight, cars_log))
boot_ci <- quantile(indirect, probs=c(0.025, 0.975))
```
**(ii) Show a density plot of the distribution of the 95% CI of the indirect effect**
```{r}
plot(density(indirect), lwd=2, col="cornflowerblue",
        main= "Distribution of the 95% CI of the indirect effect")
abline(v=quantile(indirect, probs=c(0.025, 0.975)))
```