---
title: "PBA Assignment 1"
author: "Mercy Angelina - 109006278"
date: "2022-10-26"
output: pdf_document
geometry: margin=0.4in
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
  \usepackage{booktabs}
---

# Solution: The Nuts and Bolts of Data Analysis Using R


## a) Find out the path to the directory containing the data and load it onto R

```{r}
#Read the CSV file
online_retail <- read.csv("/Users/riana.KOESNADI/Downloads/online_retail.csv", header=TRUE)
library(knitr)
kable(head(online_retail), caption = "Online Retail Data")
```

```{r}
#Check the data structure
str(online_retail)
```

## b) Convert InvoiceDate to date class and subset the data

```{r}
online_retail$NewDate <- strptime(as.character(online_retail$InvoiceDate), "%m/%d/%y")

online_retail$NewDate[261324]
class(online_retail$NewDate)

online_retail$NewInvoiceDate <- format(online_retail$NewDate, "%Y-%m-%d")

#Check InvoiceDate new format
online_retail$NewInvoiceDate[261324]

library(knitr)
kable(head(online_retail), caption = "Online Retail Data with new InvoiceDate format")

subset_df <- online_retail[online_retail$NewInvoiceDate >= "2011-07-01" & online_retail$NewInvoiceDate <= "2011-08-31", ]
subset_length <- length(subset_df$InvoiceNo)
```

* Check to see 3,664 unique transactions

```{r}
length(unique(subset_df$InvoiceNo))
```


## c) Use for-loops to
### 1) Compute the mean of Quantity and UnitPrice

```{r}
quantity_average = 0
unit_average = 0
for (i in 1: subset_length) {
  quantity_average <- quantity_average + subset_df$Quantity[i]
  unit_average <- unit_average + subset_df$UnitPrice[i]
}
quantity_average = quantity_average/subset_length
unit_average = unit_average/subset_length
```

* Mean of Quantity
```{r}
print(quantity_average)
print(unit_average)
```

### 2) Determine types of each column

```{r}
for (i in 1: length(subset_df)) {
  print(class(subset_df[,i]))
}
```

### 3) Compute the number of unique values in each column

```{r}
for (i in 1: length(subset_df)) {
  print(length(unique(subset_df[,i])))
}
```


## d) Subset the data from U.K., Netherlands, and Australia

```{r}
country_df <- subset_df[subset_df$Country == 'United Kingdom' | subset_df$Country == 'Netherlands' | subset_df$Country == 'Australia' , ]
length(unique(country_df$Country))
```

### 4) Report the average and standard deviation of UnitPrice

* Average
```{r}
x = mean(country_df$UnitPrice)
format(round(x, 3), nsmall = 3)
```
* Standard Deviation
```{r}
y = sd(country_df$UnitPrice)
format(round(y, 3), nsmall= 3)
```

### 5) Report the number of unique transactions in these countries

```{r}
length(unique(country_df$InvoiceNo))
```

### 6) Report how many customers residing in these countries made transactions in July and August of 2011?

```{r}
length(unique(country_df$CustomerID))-1
```


## e) Do we see any customers who made a refund?
Yes, we do see customers made refunds.

### 7) How many customers made a refund (exclude the observations without the **CustomerID**)?

```{r}
subset_df$Refund <- substr(subset_df$InvoiceNo, 1, 1)
cust_refund <- subset(subset_df, subset_df$Refund == "C")
length(unique(cust_refund$CustomerID))-1
```


## f) Some customers made purchases without logging into the e-commerce site. This would create records of transactions for which the **CustomerID** is missing (i.e. NA). Create a variable called **Sales** by multiplying the **Quantity** and the **UnitPrice**.

### 8) Calculate the total sales amount for those that are missing the **CustomerID** (i.e. NA)

```{r}
subset_df$Sales <- subset_df$Quantity*subset_df$UnitPrice
custNA <- subset(subset_df, is.na(subset_df$CustomerID) == TRUE)
sum(custNA$Sales)
```

### 9) How many transactions were made without the customers logging into the e-commerce site?

```{r}
length(unique(custNA$InvoiceNo))
```


## Extra Credit

### EC1) Create a variable containing the monthly aggregate spending for each customer

```{r}
#July spending
july_trans <- subset_df[subset_df$NewInvoiceDate >= "2011-07-01" & subset_df$NewInvoiceDate <= "2011-07-31", ]
#July aggregate spending for each customer
agg_cust_july <- tapply(X = july_trans$Sales, INDEX = july_trans$CustomerID, FUN = sum)

#August spending
aug_trans <- subset_df[subset_df$NewInvoiceDate >= "2011-08-01" & subset_df$NewInvoiceDate <= "2011-08-31", ]
#August aggregate spending for each customer
agg_cust_aug <- tapply(X = aug_trans$Sales, INDEX = aug_trans$CustomerID, FUN = sum)
```

### EC2) Report the IDs and the monthly purchase amount of the five customers who have spent the most money in July 2011

```{r}
agg_cust_july <- sort(-agg_cust_july)
head(agg_cust_july*(-1),5)
```